### 2021-05-05 draft.Rmd
some intro stuff

---
title: "First Draft"
author: "Group 10"
date: "5/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Packages
library(ggplot2)
library(plotly)
library(dplyr)
library(tidyverse)
library(caret)
```

```{r}
# bring in the data
stroke = read_csv("stroke.csv")
```

# Question and Initial Research
•	Question and background information on the data and why you are asking this question(s).  References to previous research/evidence generally would be nice to include. – You must present your question to me during office hours, either next week on  26th or the following week on the 3rd  
How can we most accurately predict if someone will have a stroke?

We found a dataset on Kaggle to use that includes records from both ischemic and hemorrhagic strokes. The criteria used to define a stroke was that a stroke occurs when a blood vessel that carries oxygen and nutrients to the brain is either blocked (ischemic) by a clot or bursts (hemorrhagic). This dataset includes 11 unique features of each patient, in addition to whether or not they had a stroke. The recorded features are:
• patient identification number
• gender
• age
• presence of hypertension 
• presence of heart disease
• marital status
• job type
• residence type
• average glucose level 
• BMI
• smoking status

Benefits
economic
want to predict because they're hard to deal with

# Exploratory Data Analysis
•	Exploratory Data Analysis – Initial summary statistics and graphs with an emphasis on variables you believe to be important for your analysis.

# Methods and Evaluation
Do evaluation with each method, not all after

•	Methods – Techniques you are using to address your question and the results of those methods.
•	Evaluation of your model – Select appropriate metrics and explain the output as it relates to your question.

## KNN

## Clustering

## Decision

## Random Forest

# Fairness Assessment
•	Fairness assessment – if necessary, should you happen to have any protected classes.

# Conclusions
•	Conclusions – What can you say about the results of the methods section as it relates to your question given the limitations to your model.

# Future Work
•	Future work – What additional analysis is needed or what limited your analysis on this project.

# References