### 2021-05-12 DecisionTreeFinalYu.Rmd
Decision Tree

---
title: "DecisionTreeFinalProject"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rio)
library(plyr)
library(tidyverse)
library(rpart)
library(psych)
library(pROC)
library(rpart.plot)
library(rattle)
library(caret)
library(C50)
library(mlbench)
install.packages("e1071")
library(e1071)
```

# Data Cleaning

```{r}
stroke <- read.csv("stroke.csv")
stroke <- na.omit(stroke)
stroke$bmi <- gsub("N/A", NA, stroke$bmi)
stroke$smoking_status <- gsub("Unknown", NA, stroke$smoking_status)
stroke <- na.omit(stroke)

# Data Cleaning Specifically for Decision Tree

stroke$bmi <- as.numeric(stroke$bmi)
stroke <- stroke %>%
  mutate(bmi_state = ifelse(bmi <= 25, "normal", "overweight"))

# https://www.cdc.gov/healthyweight/assessing/bmi/adult_bmi/index.html

stroke <- stroke %>%
  mutate(glucose_state = ifelse(avg_glucose_level <= 140, "normal", "prediabetic"))

# https://www.mayoclinic.org/diseases-conditions/diabetes/diagnosis-treatment/drc-20371451#:~:text=A%20blood%20sugar%20level%20less,mmol%2FL)%20indicates%20prediabetes.

stroke <- stroke %>%
  mutate(age_state = ifelse(age <= 38.4, "young", "old"))

# https://www.statista.com/statistics/241494/median-age-of-the-us-population/#:~:text=In%202018%2C%20the%20median%20age,United%20States%20was%2038.4%20years.

str(stroke)
stroke <- stroke[-c(1, 2, 3, 7, 9, 10, 11)]

stroke <- lapply(stroke, function(x) as.factor(x))

stroke <- as_tibble(stroke)

```
# Creation of Test Set

```{r}
set.seed(1950)
stroke_break <- createDataPartition(stroke$stroke, times = 1, p = 0.8, list = FALSE)
training_stroke <- stroke[stroke_break,]

test_stroke <- stroke[-stroke_break,]

```

# Baserate Calculation for Stroke
```{r}
mean(as.numeric(stroke$stroke))-1

# 0.0525394
```

# Building the Model
```{r}
set.seed(1950)
str(stroke)
as.numeric(stroke$stroke)
tree_stroke = rpart(stroke~.,  
                            method = "class",
                            parms = list(split = "gini"),
                            data = training_stroke,
                            control = rpart.control(cp=0))

```

```{r}
tree_stroke
tree_stroke$variable.importance
```

```{r}
rpart.plot(tree_stroke, type = 4, extra = 101)
plotcp(tree_stroke)

stroke_predict = predict(tree_stroke, test_stroke, type = "class")
```

```{r}
tree_matrix <- confusionMatrix(as.factor(stroke_predict), as.factor(test_stroke$stroke), positive = "1", dnn = c("Prediction", "Actual"), mode = "sens_spec")

tree_matrix
```