### 2021-05-12 KNNFinalYu.Rmd

---
title: "KNNFinalYu"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DT)
library(caret)
library(class)
```

```{r}
stroke <- read.csv("stroke.csv")
stroke$bmi <- gsub("N/A", NA, stroke$bmi)
stroke$smoking_status <- gsub("Unknown", NA, stroke$smoking_status)
stroke <- na.omit(stroke)
stroke <- stroke[ -c(1)]
str(stroke)

stroke$gender <- as.factor(stroke$gender)
stroke$ever_married <- as.factor(stroke$ever_married)
stroke$work_type <- as.factor(stroke$work_type)
stroke$Residence_type <- as.factor(stroke$Residence_type)
stroke$smoking_status <- as.factor(stroke$smoking_status)

stroke_positive <- subset(stroke, stroke == 1)

stroke_negative <- subset(stroke, stroke == 0)
set.seed(1980)
stroke_sample <- stroke_negative[sample(1:nrow(stroke_negative), size = 500),]
stroke_final <- rbind(stroke_sample, stroke_positive)

stroke <- lapply(stroke_final, function(x) as.numeric(x))

stroke <- as_tibble(stroke)

```
# Correlation Test

```{r}
stroke_correlations <- cor(stroke)
datatable(stroke_correlations)

# No columns with correlation > 0.6 or less than -0.6
```

# Creating Test and Training Set
```{r}
set.seed(1980)
stroke_break <- createDataPartition(stroke$stroke, times = 1, p = 0.8, list = FALSE)
training_stroke <- stroke[stroke_break,]

test_stroke <- stroke[-stroke_break,]
```

```{r}
chooseK = function(k, train_set, val_set, train_class, val_class){
  set.seed(1)
  class_knn = knn(train = train_set,
                  test = val_set,
                  cl = train_class,
                  k = k,
                  use.all = T)
  conf_mat = table(class_knn, val_class)
  accu = sum(conf_mat[row(conf_mat) == col(conf_mat)]) / sum(conf_mat)
  cbind(k = k, accuracy = accu)
}

different_k = sapply(seq(1, 21, by = 2),
                     function(x) chooseK(x,
                                         train_set = training_stroke[, -c(11)],
                                         val_set = test_stroke[, -c(11)],
                                         train_class = training_stroke$stroke,
                                         val_class = test_stroke$stroke
                                         ))

different_k  = tibble(k = different_k[1,],
                      recall_sen = different_k[2,])

ggplot(different_k,
       aes(x = k, y = recall_sen)) + 
  geom_line(color = "orange", size = 1.5) + 
  geom_point(size = 3)

```
# KNN Model with 3NN
```{r}
training_stroke
stroke_KNN <- knn(train = training_stroke[, -c(11)],
              test = test_stroke[, -c(11)],
              cl = training_stroke$stroke,
              k = 3,
              use.all = T,
              prob = T)

prb<- data.frame(prob = attr(stroke_KNN, "prob"))
```

```{r}
stroke_matrix <- confusionMatrix(as.factor(stroke_KNN), as.factor(test_stroke$stroke), positive = "1", dnn = c("Prediction", "Actual"), mode = "sens_spec")

stroke_matrix

```